---
title: "Restaurante and consumer data"
output: html_notebook
---

## ECDE - 2nd Semestre 2017/18

<b>Group 2:</b><br />
Alessandro Fragalà N 84080<br />
Anderson Acosta N 81799<br />
Maria Sales N 83748<br />

## Business Undestanding

The selected dataset for this project is ["Restaurante and consumer data"](https://archive.ics.uci.edu/ml/datasets/Restaurant+%26+consumer+data#), which was extracted from the UCI Machine Learning Repository.
This dataset contains information about consumers, restaurants and ratings that consumers gave to restaurants, distributed among nine .csv files: five of them about restaurants, three have consumer information and the last one contains the ratings.
The main purpose for this project is to analyze the given data and predict the missing ratings.
This project follows the CRISP-DM methodology, and makes use of the R language.

## Data Understanding

#### Library installation

Install necessary libraries:

```{r echo=TRUE}
install.packages("rminer")
library(rminer)
library(data.table)
```

The first step to start analyzing the provided data is to read all the .csv files and store them in separate datasets.

```{r echo=TRUE}
# Read datasets
# Restaurants
chefmozaccepts_df <- read.csv("chefmozaccepts.csv")
chefmozcuisine_df <- read.csv("chefmozcuisine.csv")
chefmozhours4_df <- read.csv("chefmozhours4.csv")
chefmozparking_df <- read.csv("chefmozparking.csv")
geoplaces2_df <- read.csv("geoplaces2.csv",encoding='latin-1')
# Consumers 
usercuisine_df <- read.csv("usercuisine.csv")
userpayment_df <- read.csv("userpayment.csv")
userprofile_df <- read.csv("userprofile.csv")
# User-Item-Rating
rating_final_df <- read.csv("rating_final.csv")
```

After importing all the necessary data, we need to analyze it by visualizing some important information and statistics:

#### Dataset 1: chefmozaccepts

This dataset contains information about the payment methods that restaurants accept. It contains the following features:
<li>placeID: corresponds to the restaurant id;</li>
<li>Rpayment: corresponds to the type of payment that the restaurant in the matching row accepts. This is a nominal feature with 12 levels.</li>
From this analysis we conclude that this dataset has no missing values, but a treatment to Rpayment is necessary to reduce the number of levels of this feature.
```{r}
# Columns
names(chefmozaccepts_df)
# Number of attributes
ncol(chefmozaccepts_df)
# Number of instances
nrow(chefmozaccepts_df)
# Summary
summary(chefmozaccepts_df, maxsum = 10)
# Head
head(chefmozaccepts_df)
# Number of nulls
sum(chefmozaccepts_df == "?")
# Number of nulls per column
colSums(chefmozaccepts_df == "?")
# Number of levels of Rpayment
sprintf("Rpayment has %d levels.", nlevels(chefmozaccepts_df$Rpayment))
# Number of restaurants
sprintf("There are %d unique placeID's.", length(unique(chefmozaccepts_df$placeID)))
```

#### Dataset 2: chefmozcuisine

This dataset contains information about the type of cuisine of the restaurants. It contains the following features:
<li>placeID: corresponds to the restaurant id;</li>
<li>Rcuisine: corresponds to the type of cuisine of the restaurant. This is a nominal feature with 59 levels.</li>
From this analysis we conclude that this dataset has no missing values, but a treatment to Rcuisine is necessary to reduce the number of levels of this feature.
```{r}
# Columns
names(chefmozcuisine_df)
# Number of attributes
ncol(chefmozcuisine_df)
# Number of instances
nrow(chefmozcuisine_df)
# Summary
summary(chefmozcuisine_df, maxsum = 10)
# Head
head(chefmozcuisine_df)
# Number of nulls
sum(chefmozcuisine_df == "?")
# Number of nulls per column
colSums(chefmozcuisine_df == "?")
# Number of levels of Rcuisine
sprintf("Rcuisine has %d levels.", nlevels(chefmozcuisine_df$Rcuisine))
# Number of restaurants
sprintf("There are %d unique placeID's.", length(unique(chefmozcuisine_df$placeID)))
```

#### Dataset 3: chefmozhours4

This dataset contains information about the hours and days that each restaurant is open/closed. It contains the following features:
<li>placeID: corresponds to the restaurant id;</li>
<li>hours: corresponds to the hours that the restaurant is open/closed. This is a nominal feature with 273 levels;</li>
<li>days: corresponds to the day of the week that the restaurant is open/closed. This is a nominal feature with 3 levels.</li>
From this analysis we conclude that this dataset has no missing values, but a treatment to hours is necessary to reduce the number of levels of this feature.
```{r}
# Columns
names(chefmozhours4_df)
# Number of attributes
ncol(chefmozhours4_df)
# Number of instances
nrow(chefmozhours4_df)
# Summary
summary(chefmozhours4_df, maxsum = 10)
# Head
head(chefmozhours4_df)
# Number of nulls
sum(chefmozhours4_df == "?")
# Number of nulls per column
colSums(chefmozhours4_df == "?")
# Number of levels of hours
sprintf("hours has %d levels.", nlevels(chefmozhours4_df$hours))
# Number of levels of days
sprintf("days has %d levels.", nlevels(chefmozhours4_df$days))
# Number of restaurants
sprintf("There are %d unique placeID's.", length(unique(chefmozhours4_df$placeID)))
```

#### Dataset 4: chefmozparking

This dataset contains information about the parking that each restaurant offers. It contains the following features:
<li>placeID: corresponds to the restaurant id;</li>
<li>parking_lot: corresponds to the type of parking. This is a nominal feature with 7 levels.</li>
From this analysis we conclude that this dataset has no missing values, and no treatment is necessary. If we decide to use the feature parking_lot and the data modeling takes too much time, we can consider reducing the levels of this column.
```{r}
# Columns
names(chefmozparking_df)
# Number of attributes
ncol(chefmozparking_df)
# Number of instances
nrow(chefmozparking_df)
# Summary
summary(chefmozparking_df, maxsum = 10)
# Head
head(chefmozparking_df)
# Number of nulls
sum(chefmozparking_df == "?")
# Number of nulls per column
colSums(chefmozparking_df == "?")
# Number of levels of parking_lot
sprintf("parking_lot has %d levels.", nlevels(chefmozparking_df$parking_lot))
# Number of restaurants
sprintf("There are %d unique placeID's.", length(unique(chefmozparking_df$placeID)))
```

##### Feature 4.1: parking_lot

```{r echo=TRUE}
# Auxiliar function: Draw barplot
drawBarplot <- function(dataframe, type, title, xlab, ylab, feature) {
  count <- c()
  for(s in type) {
    value <- length(dataframe[which(dataframe[,feature] == s),feature])
    count <- c(count, value)
  }
  bb <- barplot(count, main = title, xlab = xlab, ylab = ylab, names.arg = type, ylim = c(0, max(count) + max(count)*0.2)) # las = 2
  text(x = bb, y = count, labels = count, pos = 3)
}
```

A graphical visualization with a barplot shows the distribution of this column.
```{r echo=TRUE}
parking_lot <- unique(chefmozparking_df$parking_lot)
title <- "Total restaurants per parking lot type"
xlab <- "Parking Lot"
ylab <- "Restaurant count"
feature <- "parking_lot"
drawBarplot(chefmozparking_df, parking_lot, title, xlab, ylab, feature)
```

#### Dataset 5: geoplaces2

This dataset contains generic information about each restaurant. It contains the following features:
<li>placeID: corresponds to the restaurant id;</li>
<li>latitude and longitude: correspond to the geospatial coordinates of the restaurant. These are both numeric features;</li>
<li>the_geom_meter: geospatial variable. This is a nominal feature with 130 levels;</li>
<li>name: name of the restaurant;</li>
<li>address: address of the restaurant. This is a nominal feature with 100 levels and 27 missing values;</li>
<li>city: city of the restaurant. This is a nominal feature with 17 levels and 18 missing values;</li>
<li>state: state of the restaurant. This is a nominal feature with 13 levels and 18 missing values;</li>
<li>country: country of the restaurant. This is a nominal feature with 3 levels and 28 missing values;</li>
<li>fax: fax of the restaurant. This is a numeric feature with 130 missing values;</li>
<li>zip: zip code of the restaurant. This is a numeric feature with 35 levels and 74 missing values;</li>
<li>alcohol: indicates the type of bar service that the restaurant provides. This is a nominal feature with 3 levels;</li>
<li>smoking_area: indicates the type of smoking area that the restaurant provides. This is a nominal feature with 5 levels;</li>
<li>dress_code: dress code allowed in the restaurant. This is a nominal feature with 3 levels;</li>
<li>accessibility: indicates the accessibility that the restaurant  provides for disabled people. This is a nominal feature with 3 levels;</li>
<li>price: budget range of the restaurant. This is a nominal feature with 3 levels;</li>
<li>url: url of the restaurant. Contains 116 missing values;</li>
<li>Rambience: type of ambience of the restaurant. This is a nominal feature with 2 levels;</li>
<li>franchise: indicates if the restaurant is a franchise or not. This is a nominal feature with 2 levels;</li>
<li>area: indicates the type of the area of the restaurant. This is a nominal feature with 2 levels;</li>
<li>other_services: indicates if the restaurant offers any additional service or not. This is a nominal feature with 3 levels.</li>

From this analysis we conclude that this dataset needs the following treatments:
<li>drop the_geom_meter. The location of the restaurant will be given by the city</li>
<li>discover the missing cities with the given latitute and longitude, and then drop these coordinates</li>
<li>drop the name of the restaurant, because naming doesn't influence ratings</li>
<li>drop address, state, fax and zip, because these features don't influence ratings </li>
<li>drop country, because all restaurants are from Mexico (irrelevant feature)</li>

Additionaly, we conclude that in total there are 130 unique restaurants.

```{r}
# Columns
names(geoplaces2_df)
# Number of attributes
ncol(geoplaces2_df)
# Number of instances
nrow(geoplaces2_df)
# Summary
summary(geoplaces2_df)
# Head
head(geoplaces2_df)
# Number of nulls
sum(geoplaces2_df == "?")
# Number of nulls per column
colSums(geoplaces2_df == "?")
# Number of restaurants
sprintf("There are %d unique placeID's.", length(unique(geoplaces2_df$placeID)))
```

An interesting analysis is to try to understand if any of the given restaurant features affect the ratings. In order to make this analysis easier, we elaborate some connected scatterplots that will show a relation between each of the restaurant features with the ratings.

```{r echo=TRUE}
# Merge ratings and restaurants
ratings_restaurants <- merge(rating_final_df, geoplaces2_df, by = "placeID")

# Define auxiliar functions:

# Auxiliar function: Draw connected scatterplot
drawConnectedScatterplot <- function(total, rating, service, food, title, xlab, ylim, at, lab) {
  # Make a basic graph
  plot(rating~total, type="b", main=title, bty="l", xlab=xlab, ylab="Mean Rating", col=rgb(0.2,0.4,0.1,0.7), lwd=3, pch=17, ylim=ylim, xaxt='n')
  lines(service~total , col=rgb(0.8,0.4,0.1,0.7) , lwd=3 , pch=18 , type="b" )
  lines(food~total , col=rgb(0.6,0.3,0.2,0.7) , lwd=3 , pch=19 , type="b" )
  axis(1, at = at, lab = lab)
  
  # Add a legend
  legend("bottomright", legend = c("Combined", "Service", "Food"), col = c(rgb(0.2,0.4,0.1,0.7), rgb(0.8,0.4,0.1,0.7), rgb(0.6,0.3,0.2,0.7)), pch = c(17,18,19), bty = "n", pt.cex = 2, cex = 1.2, text.col = "black", horiz = F , inset = c(0.1, 0.1))
}

# Auxiliar function: Calculate mean rating
calculateMeanRating <- function(feature, value, rating) {
  mean_rating <- mean(ratings_restaurants[which(ratings_restaurants[, feature] == value), rating])
}
```

##### Feature 5.1: alcohol

```{r}
# Calculate mean rating per each type of alcohol for general rating
mean_no_alcohol_rating <- calculateMeanRating("alcohol", "No_Alcohol_Served", "rating")
mean_wine_rating <- calculateMeanRating("alcohol", "Wine-Beer", "rating")
mean_full_bar_rating <- calculateMeanRating("alcohol", "Full_Bar", "rating") 

# Calculate mean rating per each type of alcohol for service rating
mean_no_alcohol_service <- calculateMeanRating("alcohol", "No_Alcohol_Served", "service_rating")
mean_wine_service <- calculateMeanRating("alcohol", "Wine-Beer", "service_rating")
mean_full_bar_service <- calculateMeanRating("alcohol", "Full_Bar", "service_rating")

# Calculate mean rating per each type of alcohol for food rating
mean_no_alcohol_food <- calculateMeanRating("alcohol", "No_Alcohol_Served", "food_rating")
mean_wine_food <- calculateMeanRating("alcohol", "Wine-Beer", "food_rating")
mean_full_bar_food <- calculateMeanRating("alcohol", "Full_Bar", "food_rating")

# Define variables
total <- c(0, 1, 2)
rating <- c(mean_no_alcohol_rating, mean_wine_rating, mean_full_bar_rating)
service <- c(mean_no_alcohol_service, mean_wine_service, mean_full_bar_service)
food <- c(mean_no_alcohol_food, mean_wine_food, mean_full_bar_food)
title <- "Mean rating per alcohol"
xlab <- "Alcohol"
ylim <-c (1.0, 1.3)
lab <- c("No_Alcohol_Served", "Wine-Beer", "Full_Bar")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:2, lab)
```

##### Feature 5.2: smoking_area

```{r}
# Calculate mean rating per each type of smoking_area for general rating
mean_none_rating <- calculateMeanRating("smoking_area", "none", "rating")
mean_not_rating <- calculateMeanRating("smoking_area", "not permitted", "rating")
mean_bar_rating <- calculateMeanRating("smoking_area", "only at bar", "rating")
mean_permitted_rating <- calculateMeanRating("smoking_area", "permitted", "rating")
mean_section_rating <- calculateMeanRating("smoking_area", "section", "rating")

# Calculate mean rating per each type of smoking_area for service rating
mean_none_service <- calculateMeanRating("smoking_area", "none", "service_rating")
mean_not_service <- calculateMeanRating("smoking_area", "not permitted", "service_rating")
mean_bar_service <- calculateMeanRating("smoking_area", "only at bar", "service_rating")
mean_permitted_service <- calculateMeanRating("smoking_area", "permitted", "service_rating")
mean_section_service <- calculateMeanRating("smoking_area", "section", "service_rating")

# Calculate mean rating per each type of smoking_area for food rating
mean_none_food <- calculateMeanRating("smoking_area", "none", "food_rating")
mean_not_food <- calculateMeanRating("smoking_area", "not permitted", "food_rating")
mean_bar_food <- calculateMeanRating("smoking_area", "only at bar", "food_rating")
mean_permitted_food <- calculateMeanRating("smoking_area", "permitted", "food_rating")
mean_section_food <- calculateMeanRating("smoking_area", "section", "food_rating")

# Define variables
total <- c(0, 1, 2, 3, 4)
rating <- c(mean_none_rating, mean_not_rating, mean_bar_rating, mean_permitted_rating, mean_section_rating)
service <- c(mean_none_service, mean_not_service, mean_bar_service, mean_permitted_service, mean_section_service)
food <- c(mean_none_food, mean_not_food, mean_bar_food, mean_permitted_food, mean_section_food)
title <- "Mean rating per smoking area"
xlab <- "Smoking Area"
ylim <-c (0.8, 1.4)
lab <- c("None", "Not Permitted", "Only at bar", "Permitted", "Section")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:4, lab)
```

##### Feature 5.3: dress_code

```{r}
# Calculate mean rating per each type of dress_code for general rating
mean_casual_rating <- calculateMeanRating("dress_code", "casual", "rating")
mean_formal_rating <- calculateMeanRating("dress_code", "formal", "rating")
mean_informal_rating <- calculateMeanRating("dress_code", "informal", "rating")

# Calculate mean rating per each type of dress_code for service rating
mean_casual_service <- calculateMeanRating("dress_code", "casual", "service_rating")
mean_formal_service <- calculateMeanRating("dress_code", "formal", "service_rating")
mean_informal_service <- calculateMeanRating("dress_code", "informal", "service_rating")

# Calculate mean rating per each type of dress_code for food rating
mean_casual_food <- calculateMeanRating("dress_code", "casual", "food_rating")
mean_formal_food <- calculateMeanRating("dress_code", "formal", "food_rating")
mean_informal_food <- calculateMeanRating("dress_code", "informal", "food_rating")

# Define variables
total <- c(0, 1, 2)
rating <- c(mean_casual_rating, mean_formal_rating, mean_informal_rating)
service <- c(mean_casual_service, mean_formal_service, mean_informal_service)
food <- c(mean_casual_food, mean_formal_food, mean_informal_food)
title <- "Mean rating per dress type"
xlab <- "Dress type"
ylim <-c (0.8, 2.0)
lab <- c("Casual", "Formal", "Informal")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:2, lab)
```

##### Feature 5.4: accessibility

```{r}
# Calculate mean rating per each type of accessibility for general rating
mean_completely_rating <- calculateMeanRating("accessibility", "completely", "rating")
mean_no_accessibility_rating <- calculateMeanRating("accessibility", "no_accessibility", "rating")
mean_partially_rating <- calculateMeanRating("accessibility", "partially", "rating")

# Calculate mean rating per each type of accessibility for service rating
mean_completely_service <- calculateMeanRating("accessibility", "completely", "service_rating")
mean_no_accessibility_service <- calculateMeanRating("accessibility", "no_accessibility", "service_rating")
mean_partially_service <- calculateMeanRating("accessibility", "partially", "service_rating")

# Calculate mean rating per each type of accessibility for food rating
mean_completely_food <- calculateMeanRating("accessibility", "completely", "food_rating")
mean_no_accessibility_food <- calculateMeanRating("accessibility", "no_accessibility", "food_rating")
mean_partially_food <- calculateMeanRating("accessibility", "partially", "food_rating")

# Define variables
total <- c(0, 1, 2)
rating <- c(mean_completely_rating, mean_no_accessibility_rating, mean_partially_rating)
service <- c(mean_completely_service, mean_no_accessibility_service, mean_partially_service)
food <- c(mean_completely_food, mean_no_accessibility_food, mean_partially_food)
title <- "Mean rating per accessibily"
xlab <- "Accessibility"
ylim <-c (0.8, 1.4)
lab <- c("Completely", "No accessibility", "Partially")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:2, lab)
```

##### Feature 5.5: price

From the following chart we can conclude that medium and high price restaurants have better ratings than lower price restaurants.
```{r}
# Calculate mean rating per each type of price for general rating
mean_low_rating <- calculateMeanRating("price", "low", "rating")
mean_medium_rating <- calculateMeanRating("price", "medium", "rating")
mean_high_rating <- calculateMeanRating("price", "high", "rating")

# Calculate mean rating per each type of price for service rating
mean_low_service <- calculateMeanRating("price", "low", "service_rating")
mean_medium_service <- calculateMeanRating("price", "medium", "service_rating")
mean_high_service <- calculateMeanRating("price", "high", "service_rating")

# Calculate mean rating per each type of price for food rating
mean_low_food <- calculateMeanRating("price", "low", "food_rating")
mean_medium_food <- calculateMeanRating("price", "medium", "food_rating")
mean_high_food <- calculateMeanRating("price", "high", "food_rating")

# Define variables
total <- c(0, 1, 2)
rating <- c(mean_low_rating, mean_medium_rating, mean_high_rating)
service <- c(mean_low_service, mean_medium_service, mean_high_service)
food <- c(mean_low_food, mean_medium_food, mean_high_food)
title <- "Mean rating per price"
xlab <- "Price"
ylim <-c (0.9, 1.3)
lab <- c("low", "medium", "high")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:2, lab)
```

##### Feature 5.6: Rambience

```{r}
# Calculate mean rating per each type of Rambience for general rating
mean_familiar_rating <- calculateMeanRating("Rambience", "familiar", "rating")
mean_quiet_rating <- calculateMeanRating("Rambience", "quiet", "rating")

# Calculate mean rating per each type of Rambience for service rating
mean_familiar_service <- calculateMeanRating("Rambience", "familiar", "service_rating")
mean_quiet_service <- calculateMeanRating("Rambience", "quiet", "service_rating")

# Calculate mean rating per each type of Rambience for food rating
mean_familiar_food <- calculateMeanRating("Rambience", "familiar", "food_rating")
mean_quiet_food <- calculateMeanRating("Rambience", "quiet", "food_rating")

# Define variables
total <- c(0, 1)
rating <- c(mean_familiar_rating, mean_quiet_rating)
service <- c(mean_familiar_service, mean_quiet_service)
food <- c(mean_familiar_food, mean_quiet_food)
title <- "Mean rating per ambience"
xlab <- "Ambience"
ylim <-c (0.8, 1.3)
lab <- c("Familiar", "Quiet")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:1, lab)
```

##### Feature 5.7: franchise

```{r}
# Calculate mean rating per each type of franchise for general rating
mean_f_rating <- calculateMeanRating("franchise", "f", "rating")
mean_t_rating <- calculateMeanRating("franchise", "t", "rating")

# Calculate mean rating per each type of franchise for service rating
mean_f_service <- calculateMeanRating("franchise", "f", "service_rating")
mean_t_service <- calculateMeanRating("franchise", "t", "service_rating")

# Calculate mean rating per each type of franchise for food rating
mean_f_food <- calculateMeanRating("franchise", "f", "food_rating")
mean_t_food <- calculateMeanRating("franchise", "t", "food_rating")

# Define variables
total <- c(0, 1)
rating <- c(mean_f_rating, mean_t_rating)
service <- c(mean_f_service, mean_t_service)
food <- c(mean_f_food, mean_t_food)
title <- "Mean rating per franchise"
xlab <- "Franchise"
ylim <-c (0.8, 1.3)
lab <- c("No", "Yes")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:1, lab)
```

##### Feature 5.8: area

```{r}
# Calculate mean rating per each type of area for general rating
mean_closed_rating <- calculateMeanRating("area", "closed", "rating")
mean_open_rating <- calculateMeanRating("area", "open", "rating")

# Calculate mean rating per each type of area for service rating
mean_closed_service <- calculateMeanRating("area", "closed", "service_rating")
mean_open_service <- calculateMeanRating("area", "open", "service_rating")

# Calculate mean rating per each type of area for food rating
mean_closed_food <- calculateMeanRating("area", "closed", "food_rating")
mean_open_food <- calculateMeanRating("area", "open", "food_rating")

# Define variables
total <- c(0, 1)
rating <- c(mean_closed_rating, mean_open_rating)
service <- c(mean_closed_service, mean_open_service)
food <- c(mean_closed_food, mean_open_food)
title <- "Mean rating per area"
xlab <- "Area"
ylim <-c (0.8, 1.3)
lab <- c("Closed", "Oper")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:1, lab)
```

##### Feature 5.9: other_services

```{r}
# Calculate mean rating per each type of other_services for general rating
mean_Internet_rating <- calculateMeanRating("other_services", "Internet", "rating")
mean_none_rating <- calculateMeanRating("other_services", "none", "rating")
mean_variety_rating <- calculateMeanRating("other_services", "variety", "rating")

# Calculate mean rating per each type of other_services for service rating
mean_Internet_service <- calculateMeanRating("other_services", "Internet", "service_rating")
mean_none_service <- calculateMeanRating("other_services", "none", "service_rating")
mean_variety_service <- calculateMeanRating("other_services", "variety", "service_rating")

# Calculate mean rating per each type of other_services for food rating
mean_Internet_food <- calculateMeanRating("other_services", "Internet", "food_rating")
mean_nonefood <- calculateMeanRating("other_services", "none", "food_rating")
mean_variety_food <- calculateMeanRating("other_services", "variety", "food_rating")

# Define variables
total <- c(0, 1, 2)
rating <- c(mean_Internet_rating, mean_none_rating, mean_variety_rating)
service <- c(mean_Internet_service, mean_none_service, mean_variety_service)
food <- c(mean_Internet_food, mean_none_food, mean_variety_food)
title <- "Mean rating per other services"
xlab <- "Other service"
ylim <-c (1.0, 1.6)
lab <- c("Internet", "None", "Variety")

# Draw graph
drawConnectedScatterplot(total, rating, service, food, title, xlab, ylim, 0:2, lab)

```

#### Dataset 6: usercuisine

This dataset contains information about the preferred cuisine(s) type(s) for each user. It contains the following features:
<li>userID: corresponds to the user id;</li>
<li>Rcuisine: corresponds to the type(s) of cuisine(s) that the user prefers. This is a nominal feature with 103 levels.</li>
From this analysis we conclude that this dataset has no missing values, but a treatment to Rcuisine is necessary to reduce the number of levels of this feature. This treatment should match the one done to Rcuisine from chefmozcuisine.
```{r}
# Columns
names(usercuisine_df)
# Number of attributes
ncol(usercuisine_df)
# Number of instances
nrow(usercuisine_df)
# Summary
summary(usercuisine_df)
# Head
head(usercuisine_df)
# Number of nulls
sum(usercuisine_df == "?")
# Number of nulls per column
colSums(usercuisine_df == "?")
# Number of levels of Rcuisine
sprintf("Rcuisine has %d levels.", nlevels(usercuisine_df$Rcuisine))
# Number of users
sprintf("There are %d unique userID's.", length(unique(usercuisine_df$userID)))
```

#### Dataset 7: userpayment

This dataset contains information about the payment methods that users prefer. It contains the following features:
<li>userID: corresponds to the user id;</li>
<li>Upayment: corresponds to the type of payment that the user prefers. This is a nominal feature with 5 levels.</li>
From this analysis we conclude that this dataset has no missing values, but a treatment to Upayment is necessary to reduce the number of levels of this feature, and it should match the treatment made to chefmozaccepts.
```{r}
# Columns
names(userpayment_df)
# Number of attributes
ncol(userpayment_df)
# Number of instances
nrow(userpayment_df)
# Summary
summary(userpayment_df)
# Head
head(userpayment_df)
# Number of nulls
sum(userpayment_df == "?")
# Number of nulls per column
colSums(userpayment_df == "?")
# Number of levels of Rpayment
sprintf("Upayment has %d levels.", nlevels(userpayment_df$Upayment))
# Number of users
sprintf("There are %d unique userID's.", length(unique(userpayment_df$userID)))
```

##### Feature 7.1: Upayment

A graphical visualization with a barplot shows the distribution of this feature.
```{r}
upayment <- unique(userpayment_df$Upayment)
title <- "Total users per payment type"
xlab <- "UPayment"
ylab <- "User count"
feature <- "Upayment"
drawBarplot(userpayment_df, upayment, title, xlab, ylab, feature)
```

#### Dataset 8: userprofile

This dataset contains information about the users that rated the restaurants. It contains the following features:
<li>userID: corresponds to the user id;</li>
<li>latitude and longitude: correspond to the geospatial coordinates of the user These are both numeric features;</li>
<li>smoker: indicates whether the user is smoker or not. This is a nominal feature with 2 levels and 3 missing values;</li>
<li>drink_level: indicates the drink level of the user. This is a nominal feature with 3 levels;</li>
<li>dress_preference: indicates the user's dress preference. This is a nominal feature with 4 levels and 5 missing values;</li>
<li>ambience: indicates the ambience that the user prefers. This is a nominal feature with 3 levels and 6 missing values;</li>
<li>transport: indicates the kind of transportation that the user prefers. This is a nominal feature with 3 levels and 7 missing values;</li>
<li>marital_status: indicates the marital status of the user. This is a nominal feature with 3 levels and 4 missing values;</li>
<li>hijos: indicates wheter the user has kids or not. This is a nominal feature with 3 levels and 11 missing values;</li>
<li>birth_year: user birth year. This is a numeric feature;</li>
<li>interest: indicates user's interest. This is a nominal feature with 5 levels;</li>
<li>personality: indicates user's personality. This is a nominal feature with 4 levels;</li>
<li>religion: indicates user's religion. This is a nominal feature with 5 levels;</li>
<li>activity: indicates user's activity. This is a nominal feature with 4 levels and 7 missing values;</li>
<li>color: indicates user's color preference. This is a nominal feature with 8 levels;</li>
<li>weight: user's weight (numeric);</li>
<li>budget: indicates user's budget. This is a nominal feature with 3 levels and 7 missing values;</li>
<li>height: user's height (numeric).</li>

From this analysis we conclude that this dataset needs the following treatments:
<li>discover the user cities to match the restaurant cities with the given latitute and longitude, and then drop these coordinates</li>
<li>treat smoker missing values</li>
<li>treat dress_preference missing values</li>
<li>treat ambience missing values</li>
<li>treat transport missing values</li>
<li>treat marital_status missing values</li>
<li>treat hijos missing values</li>
<li>treat activity missing values</li>
<li>treat budget missing values</li>

Additionaly, we conclude that in total there are 138 unique users.
```{r}
# Columns
names(userprofile_df)
# Number of attributes
ncol(userprofile_df)
# Number of instances
nrow(userprofile_df)
# Summary
summary(userprofile_df)
# Head
head(userprofile_df)
# Number of nulls
sum(userprofile_df == "?")
# Number of nulls per column
colSums(userprofile_df == "?")
# Number of users
sprintf("There are %d unique userID's.", length(unique(userprofile_df$userID)))
```

Graphical visualization:

##### Feature 8.1: smoker
```{r}
smoker_type <- unique(levels(userprofile_df$smoker))
title <- "Total of users by smoker type"
xlab <- "Smoker type"
ylab <- "User count"
feature <- "smoker"
drawBarplot(userprofile_df, smoker_type, title, xlab, ylab, feature)
```

##### Feature 8.2: drink_level
```{r}
drink_level_type <- unique(levels(userprofile_df$drink_level))
title <- "Total users per drink level"
xlab <- "Drink level"
ylab <- "User count"
feature <- "drink_level"
drawBarplot(userprofile_df, drink_level_type, title, xlab, ylab, feature)
```

##### Feature 8.3: dress_preference
```{r}
dress_preference_type <- unique(levels(userprofile_df$dress_preference))
title <- "Total users per dress preference"
xlab <- "Dress Preference"
ylab <- "User count"
feature <- "dress_preference"
drawBarplot(userprofile_df, dress_preference_type, title, xlab, ylab, feature)
```

##### Feature 8.4: ambience
```{r}
ambience_type <- unique(levels(userprofile_df$ambience))
title <- "Total users per ambience type"
xlab <- "Ambience"
ylab <- "User count"
feature <- "ambience"
drawBarplot(userprofile_df, ambience_type, title, xlab, ylab, feature)
```

##### Feature 8.5: transport
```{r}
transport_type <- unique(levels(userprofile_df$transport))
title <- "Total users per transport type"
xlab <- "Transport"
ylab <- "User count"
feature <- "transport"
drawBarplot(userprofile_df, transport_type, title, xlab, ylab, feature)
```

##### Feature 8.6: marital_status
```{r}
marital_status <- unique(levels(userprofile_df$marital_status))
title <- "Total users per marital status"
xlab <- "Marital Status"
ylab <- "User count"
feature <- "marital_status"
drawBarplot(userprofile_df, marital_status, title, xlab, ylab, feature)
```

##### Feature 8.7: hijos
```{r}
hijos <- unique(levels(userprofile_df$hijos))
title <- "Total users per children"
xlab <- "Children"
ylab <- "User count"
feature <- "hijos"
drawBarplot(userprofile_df, hijos, title, xlab, ylab, feature)
```

##### Feature 8.8: interest
```{r}
interest <- unique(levels(userprofile_df$interest))
title <- "Total users per interest"
xlab <- "Interest"
ylab <- "User count"
feature <- "interest"
drawBarplot(userprofile_df, interest, title, xlab, ylab, feature)
```

##### Feature 8.9: personality
```{r}
personality <- unique(levels(userprofile_df$personality))
title <- "Total users per personality"
xlab <- "Personality"
ylab <- "User count"
feature <- "personality"
drawBarplot(userprofile_df, personality, title, xlab, ylab, feature)
```

##### Feature 8.10: religion
```{r}
religion <- unique(levels(userprofile_df$religion))
title <- "Total users per religion"
xlab <- "Religion"
ylab <- "User count"
feature <- "religion"
drawBarplot(userprofile_df, religion, title, xlab, ylab, feature)
```

##### Feature 8.11: activity
```{r}
activity <- unique(levels(userprofile_df$activity))
title <- "Total users per activity"
xlab <- "Activity"
ylab <- "User count"
feature <- "activity"
drawBarplot(userprofile_df, activity, title, xlab, ylab, feature)
```

##### Feature 8.12: color
```{r}
color <- unique(levels(userprofile_df$color))
title <- "Total users per color"
xlab <- "Color"
ylab <- "User count"
feature <- "color"
drawBarplot(userprofile_df, color, title, xlab, ylab, feature)
```

##### Feature 8.13: budget
```{r}
budget_type <- unique(levels(userprofile_df$budget))
title <- "Total users per budget type"
xlab <- "Budget"
ylab <- "User count"
feature <- "budget"
drawBarplot(userprofile_df, budget_type, title, xlab, ylab, feature)
```

#### Dataset 9: rating_final

The final dataset contains information about the ratings that users gave to restaurants. It contains the following features:
<li>userID: corresponds to the user id;</li>
<li>placeID: corresponds to the restaurant id;</li>
<li>rating: combined rating of the restaurant. This is a numerical feature;</li>
<li>food_rating: rating that corresponds to the quality of the food of the restaurant. This is a numerical feature;</li>
<li>service_rating: rating that corresponds to the quality of the service of the restaurant. This is a numerical feature.</li>

It's interesting to notice that even though rating, food_rating and service_rating are presented as numerical features, they only have 3 possible values: 0, 1 or 2. This means that the problem of predicting the ratings can either be treated as a regression or classification problem.
```{r}
# Columns
names(rating_final_df)
# Number of attributes
ncol(rating_final_df)
# Number of instances
nrow(rating_final_df)
# Summary
summary(rating_final_df)
# Head
head(rating_final_df)
# Number of nulls
sum(rating_final_df == "?")
# Number of nulls per column
colSums(rating_final_df == "?")
# Number of users
sprintf("There are %d unique userID's.", length(unique(rating_final_df$userID)))
# Number of resstaurant
sprintf("There are %d unique placeID's.", length(unique(rating_final_df$placeID)))
```

#### Conclusion

In total, there are 938 restaurants and 138 users, but only 130 restaurants have ratings. Each restaurant is characterized by a set of features, and each customer preference is also described by a set of features with the same dimension.
```{r}
total_restaurants <- length(unique(c(chefmozaccepts_df$placeID, chefmozcuisine_df$placeID, chefmozhours4_df$placeID, chefmozparking_df$placeID, geoplaces2_df$placeID)))
total_users <- length(unique(c(usercuisine_df$userID, userpayment_df$userID, userprofile_df$userID)))
sprintf("There are %d restaurants and %d users.", total_restaurants, total_users)
```

## Data Preparation

#### Dataset 1: chefmozaccepts

As observed in Data Understanding, this dataset has an attribute (Rpayment) that has 12 levels. An important treatment to this feature is to reduce the amount of levels. The same problem exists in userpayment dataframe with Upayment, but instead this one has 5 levels. To accommodate the treatment for both datasets, the following levels will be created (instead of the originals presented in the datasets):

1. Card 
  - Instead of: American_Express, bank_debit_cards, Carte_Blanche, Visa, VISA, MasterCard-Eurocard
2. Cash
  - Instead of: cash
3. Other
  - Instead of: checks, Diners_Club, Discover, gift_certificates, Japan_Credit_Bureau, 

```{r echo=TRUE}
# Auxiliar function: Calculate new payment feature
calculatePayment <- function(dataset) {
  
  dataset$payment <- c("")
  
  # Compute new payment types
  for(i in seq_len(nrow(dataset))) {
    payment <- as.character(dataset[i,2])
    if(payment == "American_Express" || payment == "bank_debit_cards" || payment == "Carte_Blanche" || payment == "Visa" || payment == "VISA" || payment == "MasterCard-Eurocard") {
      dataset[i,3] <- "Card"
    } else if(payment == "cash") {
      dataset[i,3] <- "Cash"
    } else {
      dataset[i,3] <- "Other"
    }
  }
  
  # Transform Rpayment to calculated values as factor
  dataset[,2] <- as.factor(dataset$payment)
  
  # Drop payment column
  drops <- c("payment")
  dataset <- dataset[ , !(names(dataset) %in% drops)]
  
  # Return only unique rows
  return(unique(dataset))
}
```

```{r}
# Calculate new Rpayment types
chefmozaccepts_df <- calculatePayment(chefmozaccepts_df)

summary(chefmozaccepts_df$Rpayment)
```

#### Datset 2: chefmozcuisine

As observed in Data Understanding, this dataset has an attribute (Rcuisine) that has 59 levels. An important treatment to this feature is to reduce the amount of levels. The same problem exists in usercuisine dataframe, but instead this one has 103 levels. To accommodate the treatment for both datasets, the following levels will be created (instead of the originals presented in the datasets):

1. Bar/Cafeteria 
  - Instead of: Bagels, Bakery, Bar, Bar_Pub_Brewery, Breakfast-Brunch, Cafe-Coffee_Shop, Cafeteria, Deli-Sandwiches, Dessert-Ice_Cream, Doughnuts, Juice
2. Mexican
  - Instead of: Mexican, Regional, Tex-Mex
3. Asian
  - Instead of: Asian, Cambodian, Chinese, Dim_Sum, Filipino, Japanese, Korean, Malaysian, Mongolian, Sushi, Thai, Tibetan, Vietnamese
4. Internationl (others)
  - Instead of: Afghan, African, American, Armenian, Barbecue, Brazilian, Burgers, California, Caribbean, Contemporary, Continental-European, Diner, Dutch-Belgian, Eastern_European, Ethiopian, Family, Fast_Food, Fine_Dining, French, Game, German, Greek, Hot_Dogs, International, Italian, Latin_American, Mediterranean, Organic-Healthy, Persian, Pizzeria, Polish, Seafood, Soup, Southern, Southwestern, Spanish, Steaks, Turkish, Vegetarian

```{r}
# Auxiliar function: Calculate new cuisine feature
calculateCuisine <- function(dataset) {
  
  dataset$cuisine <- c("")
  
  # Compute new cuisine types
  for(i in seq_len(nrow(dataset))) {
    cuisine <- as.character(dataset[i,2])
    if(cuisine == "Bagels" || cuisine == "Bakery" || cuisine == "Bar" || cuisine == "Bar_Pub_Brewery" || cuisine == "Breakfast-Brunch" || cuisine == "Cafe-Coffee_Shop" || cuisine == "Cafeteria" || cuisine == "Deli-Sandwiches" || cuisine == "Dessert-Ice_Cream" || cuisine == "Doughnuts" || cuisine == "Juice") {
      dataset[i,3] <- "Bar/Cafeteria"
    } else if(cuisine == "Mexican" || cuisine == "Regional" || cuisine == "Tex-Mex") {
      dataset[i,3] <- "Mexican"
    } else if(cuisine == "Asian" || cuisine == "Cambodian" || cuisine == "Chinese" || cuisine == "Dim_Sum" || cuisine == "Filipino" || cuisine == "Japanese" || cuisine == "Korean" || cuisine == "Malaysian" || cuisine == "Mongolian" || cuisine == "Sushi" || cuisine == "Thai" || cuisine == "Tibetan" || cuisine == "Vietnamese") {
      dataset[i,3] <- "Asian"
    } else {
      dataset[i,3] <- "International"
    }
  }
  
  # Transform Rcuisine to calculated values as factor
  dataset[,2] <- as.factor(dataset$cuisine)
  
  # Drop cuisine column
  drops <- c("cuisine")
  dataset <- dataset[ , !(names(dataset) %in% drops)]
  
  # Return only unique rows
  return(unique(dataset))
}
```

```{r}
# Calculate new Rcuisine types
chefmozcuisine_df <- calculateCuisine(chefmozcuisine_df)

summary(chefmozcuisine_df$Rcuisine)
```

#### Dataset 3: chefmozhours4

// TODO

#### Dataset 4: chefmozparking

As discussed in Data Undestarnding, no treatment will be applied to this dataset.

#### Dataset 5: geoplaces2

Select only relevant features.
Treat missing hours and create relevant features.

```{r}
rest_df <- geoplaces2_df[,c("placeID", "latitude", "longitude", "city", "alcohol", 
                            "smoking_area", "dress_code", "accessibility", "price", "Rambience",
                            "franchise", "area", "other_services")]

rest_st_df <- chefmozhours4_df[which(chefmozhours4_df$days == "Sat;"),]

rest_sd_df <- chefmozhours4_df[which(chefmozhours4_df$days == "Sun;"),]
rest_wk_df <- chefmozhours4_df[which(chefmozhours4_df$days == "Mon;Tue;Wed;Thu;Fri;"),]

rest_days_df <- merge(rest_st_df, rest_sd_df, by = "placeID")
rest_days_df <- merge(rest_days_df, rest_wk_df, by = "placeID")

setnames(rest_days_df, old=c("hours.x","hours.y","hours"), new=c("st", "sd","wk"))

rest_days_df <- rest_days_df[,c("placeID", "st", "sd","wk")]

rest_days_df <- rest_days_df[!duplicated(rest_days_df[c("placeID", "st", "sd","wk")]),]

rest_days_df <- data.frame(lapply(rest_days_df, as.character), stringsAsFactors = FALSE)
  
restaurant_df <- merge(rest_df, rest_days_df, by = "placeID", all.x = TRUE)

for(placeID in unique(restaurant_df$placeID)) {
  test <- restaurant_df[which(restaurant_df$placeID == placeID),]
  if(is.element("00:00-00:00;", test$st)){
    #Closed restaurant
    restaurant_df[which(restaurant_df$placeID == placeID),]$st <- 0
  } else {
    #Open restaurant (the Na will considered "open")
    restaurant_df[which(restaurant_df$placeID == placeID),]$st <- 1
  }
}

for(placeID in unique(restaurant_df$placeID)) {
  test <- restaurant_df[which(restaurant_df$placeID == placeID),]
  if(is.element("00:00-00:00;", test$sd)){
    #Closed restaurant
    restaurant_df[which(restaurant_df$placeID == placeID),]$sd <- 0
  } else {
    #Open restaurant (the Na will considered "open")
    restaurant_df[which(restaurant_df$placeID == placeID),]$sd <- 1
  }
}

for(placeID in unique(restaurant_df$placeID)) {
  test <- restaurant_df[which(restaurant_df$placeID == placeID),]
  if(is.element("00:00-00:00;", test$st)){
    #Closed restaurant
    restaurant_df[which(restaurant_df$placeID == placeID),]$wk <- 0
  } else {
    #Open restaurant (the Na will considered "open")
    restaurant_df[which(restaurant_df$placeID == placeID),]$wk <- 1
  }
}

geoplaces2_df <- restaurant_df

summary(geoplaces2_df)
```

#### Dataset 6: usercuisine

As discussed in Data Understanding section, a treatment to Ucuisine will be applied to reduce the number of levels. This treatment matches the one perfomed to chefmozcuisine dataset.

```{r}
# Calculate new RCuisine types
usercuisine_df <- calculateCuisine(usercuisine_df)

summary(usercuisine_df$Rcuisine)

# Rename Rcuisine to Ucuisine
names(usercuisine_df) <- c("userID", "Ucuisine")
```

#### Dataset 7: userpayment

As discussed in Data Understanding section, a treatment to Upayment will be applied to reduce the number of levels of this feature. This treatment matches the one performed to chefmozaccepts dataset.

```{r}
# Calculate new Upayment types
userpayment_df <- calculatePayment(userpayment_df)

summary(userpayment_df$Upayment)
```

#### Dataset 8: userprofile

As seen in Data Understanding phase, userprofile dataset has 50 missing values. The columns with missing values are 'smoker', 'ambience', 'transport', 'marital_status', 'hijos', 'activity' and 'budget'. These are all categorical features, so we decided to replace the missing values with the corresponding mode of that column.

```{r echo=TRUE}
# Auxiliar function: Calculate mode
calculateMode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

##### Feature 8.1: smoker

This column has 50 missing values, where 26 are 'true' (smoker) and 112 are 'false' (non smoker). This means that the mode for this column is 'false'. In this case, and because it's preferrable to put a smoker person in a non-smoking restaurant than a non-smoker in a smoker restaurant, we will impute the missing values with 'false' (non smoker).

```{r}
summary(userprofile_df$smoker)

# Force NA's
userprofile_df[which(userprofile_df$smoker=="?"),]$smoker <- NA
length(which(is.na(userprofile_df$smoker)))

# Calculate mode
smoker_mode <- calculateMode(na.omit(userprofile_df$smoker))
sprintf("Mode: %s", smoker_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "smoker", Value=smoker_mode)
print(summary(userprofile_df$smoker))
```

##### Feature 8.3: dress_preference

This column has 5 missing values. The mode for this column is 'no preference'.
```{r}
summary(userprofile_df$dress_preference)

# Force NA's
userprofile_df[which(userprofile_df$dress_preference=="?"),]$dress_preference <- NA
length(which(is.na(userprofile_df$dress_preference)))

# Calculate mode
dress_preference_mode <- calculateMode(na.omit(userprofile_df$dress_preference))
sprintf("Mode: %s", dress_preference_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "dress_preference", Value=dress_preference_mode)
print(summary(userprofile_df$dress_preference))
```

##### Feature 8.4: ambience

This column has 6 missing values. The mode for this column is 'family'.

```{r}
summary(userprofile_df$ambience)

# Force NA's
userprofile_df[which(userprofile_df$ambience=="?"),]$ambience <- NA
length(which(is.na(userprofile_df$ambience)))

# Calculate mode
ambience_mode <- calculateMode(na.omit(userprofile_df$ambience))
sprintf("Mode: %s", ambience_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "ambience", Value=ambience_mode)
print(summary(userprofile_df$ambience))
```

##### Feature 8.5: transport

This column has 7 missing values, and the mode is 'public'.
```{r}
summary(userprofile_df$transport_mode)

# Force NA's
userprofile_df[which(userprofile_df$transport=="?"),]$transport <- NA
length(which(is.na(userprofile_df$transport)))

# Calculate mode
transport_mode <- calculateMode(na.omit(userprofile_df$transport))
sprintf("Mode: %s", transport_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "transport", Value=transport_mode)
print(summary(userprofile_df$transport))
```

##### Feature 8.6: marital_status

This column has 4 missing values, and the mode is 'single'.
```{r}
summary(userprofile_df$marital_status)

# Force NA's
userprofile_df[which(userprofile_df$marital_status=="?"),]$marital_status <- NA
length(which(is.na(userprofile_df$marital_status)))

# Calculate mode
marital_status_mode <- calculateMode(na.omit(userprofile_df$marital_status))
sprintf("Mode: %s", marital_status_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "marital_status", Value=marital_status_mode)
print(summary(userprofile_df$marital_status))
```

##### Feature 8.7: hijos

This column has 11 missing values, and the mode is 'independent'.
```{r}
summary(userprofile_df$hijos)

# Force NA's
userprofile_df[which(userprofile_df$hijos=="?"),]$hijos <- NA
length(which(is.na(userprofile_df$hijos)))

# Calculate mode
hijos_mode <- calculateMode(na.omit(userprofile_df$hijos))
sprintf("Mode: %s", hijos_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "hijos", Value=hijos_mode)
print(summary(userprofile_df$hijos))
```

##### Feature 8.11: activity

This column has 7 missing values, and the mode is 'student'.
```{r}
summary(userprofile_df$activity)

# Force NA's
userprofile_df[which(userprofile_df$activity=="?"),]$activity <- NA
length(which(is.na(userprofile_df$activity)))

# Calculate mode
activity_mode <- calculateMode(na.omit(userprofile_df$activity))
sprintf("Mode: %s", activity_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "activity", Value=activity_mode)
print(summary(userprofile_df$activity))
```

##### Feature 8.13: budget

This column has 7 missing values, and the mode is 'medium'.
```{r}
summary(userprofile_df$budget)

# Force NA's
userprofile_df[which(userprofile_df$budget=="?"),]$budget <- NA
length(which(is.na(userprofile_df$budget)))

# Calculate mode
budget_mode <- calculateMode(na.omit(userprofile_df$budget))
sprintf("Mode: %s", budget_mode)

# Substitute
userprofile_df <- imputation("value", userprofile_df, "budget", Value=budget_mode)
print(summary(userprofile_df$budget))
```


#### Dataset 9: rating_final

As discussed in Data Understanding, this dataset doesn't need any treatment because it contains the ratings that users gave to the restaurants.

The dataset that will be used for the modelling part should include all the treatments applied during this phase and the relevant features that were identified.

```{r}
# Join dataframes
final_df <- merge(rating_final_df, chefmozaccepts_df, by = "placeID")
final_df <- merge(final_df, chefmozcuisine_df, by = "placeID")
final_df <- merge(final_df, chefmozhours4_df, by = "placeID")
final_df <- merge(final_df, chefmozparking_df, by = "placeID")
final_df <- merge(final_df, geoplaces2_df, by = "placeID")
final_df <- merge(final_df, usercuisine_df, by = "userID")
final_df <- merge(final_df, userpayment_df, by = "userID")
final_df <- merge(final_df, userprofile_df, by = "userID")

# Rename a column in R
colnames(final_df)[colnames(final_df)=="Rcuisine.x"] <- "Rcuisine"
colnames(final_df)[colnames(final_df)=="Rcuisine.y"] <- "UserRcuisine"

# Analyze final dataframe
# Columns
names(final_df)
# Number of attributes
ncol(final_df)
# Number of instances
nrow(final_df)
# Summary
summary(final_df)
# Head
head(final_df)
# Number of nulls
sum(final_df == "?")
# Number of nulls per column
colSums(final_df == "?")

# Drop unnecessary columns
final_df <- final_df[,c("rating","Rpayment", "days", "parking_lot", "Upayment", "smoker", "drink_level", "dress_preference", "ambience", "transport", "marital_status", "hijos", "interest", "personality", "religion", "activity", "color", "budget")]

# Dropped columns: userID, placeID, food_rating, service_rating, latitude, longitude, birth_year, weight, height, hours, UserRcuisine, Rcuisine
```

## Modeling

The main goal for this project is to predict the ratings that would be given by each customer for the restaurants that he/she didn't rate. In Data Understanding we saw that ratings are numerical values, so predicting their values can be treated as a regression problem. Using classification techniques is also reasonable since the possible ratings are three discrete integers: 0, 1 and 2. We will focus on predicting the ratings, and not predict food_rating or service_rating.


### Holdout
As this is a supervised learning problem, the first step is to select the training and test datasets.
We will use the holdout technique.

```{r}

summary(final_df)
H = holdout(final_df$rating, ratio = 0.8)
```

### Neuronal Network

```{r}
?fit
Fi <- fit(rating~., final_df[H$tr,], model = "mlp")

P <- predict(Fi, final_df[H$ts,])

mmetric(final_df[H$ts,]$rating, P, metric = "MAE")
mmetric(final_df[H$ts,]$rating, P, metric = "MAPE")
```

## Evaluation

## Deployment

This step will not be executed.

# Bibliography

https://www.kaggle.com/liyenhsu/simple-content-based-recommenders

Get mode: https://www.tutorialspoint.com/r/r_mean_median_mode.htm





