---
title: "Restaurante and consumer data"
output: html_notebook
---

## ECDE - 2nd Semestre 2017/18

<b>Group 2:</b><br />
Alessandro Fragalà N 84080<br />
Anderson Acosta N 81799<br />
Maria Sales N 83748<br />

The selected dataset for this project is ["Restaurante and consumer data"](https://archive.ics.uci.edu/ml/datasets/Restaurant+%26+consumer+data#), which was extracted from the UCI Machine Learning Repository.
This dataset contains nine csv files: five of them have Restaurant categories, three have consumer categories and the last one contains ratings.

## Business Undestanding

The main purpose for this project is to "generate a top-n list of restaurants according to the consumer preferences". 

## Data Understanding

First step is to read all the datasets
```{r}
# Set workspace
setwd("/Users/mariasales/Desktop/MIG/1A2A/ECDE/Trabalho")

# Read datasets
#Restaurants
chefmozaccepts_df <- read.csv("chefmozaccepts.csv")
chefmozcuisine_df <- read.csv("chefmozcuisine.csv")
chefmozhours4_df <- read.csv("chefmozhours4.csv")
chefmozparking_df <- read.csv("chefmozparking.csv")
geoplaces2_df <- read.csv("geoplaces2.csv",encoding='latin-1')
#Consumers 
usercuisine_df <- read.csv("usercuisine.csv")
userpayment_df <- read.csv("userpayment.csv")
userprofile_df <- read.csv("userprofile.csv")
#User-Item-Rating
rating_final_df <- read.csv("rating_final.csv")
```

Then we want to analyze the imported data, as see some important statistics about all files:

#### chefmozaccepts.csv
```{r}
# Columns
names(chefmozaccepts_df)
# Number of attributes
ncol(chefmozaccepts_df)
# Number of instances
nrow(chefmozaccepts_df)
# Summary
summary(chefmozaccepts_df, maxsum = 1000)
# Head
head(chefmozaccepts_df)
# Number of nulls
sum(chefmozaccepts_df == "?")
# Number of nulls per column
colSums(chefmozaccepts_df == "?")
```

#### chefmozcuisine.csv
```{r}
# Columns
names(chefmozcuisine_df)
# Number of attributes
ncol(chefmozcuisine_df)
# Number of instances
nrow(chefmozcuisine_df)
# Summary
summary(chefmozcuisine_df, maxsum = 1000)
# Head
head(chefmozcuisine_df)
# Number of nulls
sum(chefmozcuisine_df == "?")
# Number of nulls per column
colSums(chefmozcuisine_df == "?")
```

#### chefmozhours4.csv
```{r}
# Columns
names(chefmozhours4_df)
# Number of attributes
ncol(chefmozhours4_df)
# Number of instances
nrow(chefmozhours4_df)
# Summary
summary(chefmozhours4_df, maxsum = 1000)
# Head
head(chefmozhours4_df)
# Number of nulls
sum(chefmozhours4_df == "?")
# Number of nulls per column
colSums(chefmozhours4_df == "?")
```

#### chefmozparking_df
```{r}
# Columns
names(chefmozparking_df)
# Number of attributes
ncol(chefmozparking_df)
# Number of instances
nrow(chefmozparking_df)
# Summary
summary(chefmozparking_df, maxsum = 1000)
# Head
head(chefmozparking_df)
# Number of nulls
sum(chefmozparking_df == "?")
# Number of nulls per column
colSums(chefmozparking_df == "?")
```

#### geoplaces2_df
```{r}
# Columns
names(geoplaces2_df)
# Number of attributes
ncol(geoplaces2_df)
# Number of instances
nrow(geoplaces2_df)
# Summary
summary(geoplaces2_df, maxsum = 1000)
# Head
head(geoplaces2_df)
# Number of nulls
sum(geoplaces2_df == "?")
# Number of nulls per column
colSums(geoplaces2_df == "?")
```

#### usercuisine_df
```{r}
# Columns
names(usercuisine_df)
# Number of attributes
ncol(usercuisine_df)
# Number of instances
nrow(usercuisine_df)
# Summary
summary(usercuisine_df, maxsum = 1000)
# Head
head(usercuisine_df)
# Number of nulls
sum(usercuisine_df == "?")
# Number of nulls per column
colSums(usercuisine_df == "?")
```

#### userpayment_df
```{r}
# Columns
names(userpayment_df)
# Number of attributes
ncol(userpayment_df)
# Number of instances
nrow(userpayment_df)
# Summary
summary(userpayment_df, maxsum = 1000)
# Head
head(userpayment_df)
# Number of nulls
sum(userpayment_df == "?")
# Number of nulls per column
colSums(userpayment_df == "?")
```

#### userprofile_df
```{r}
# Columns
names(userprofile_df)
# Number of attributes
ncol(userprofile_df)
# Number of instances
nrow(userprofile_df)
# Summary
summary(userprofile_df, maxsum = 1000)
# Head
head(userprofile_df)
# Number of nulls
sum(userprofile_df == "?")
# Number of nulls per column
colSums(userprofile_df == "?")
```

#### rating_final_df
```{r}
# Columns
names(rating_final_df)
# Number of attributes
ncol(rating_final_df)
# Number of instances
nrow(rating_final_df)
# Summary
summary(rating_final_df, maxsum = 1000)
# Head
head(rating_final_df)
# Number of nulls
sum(rating_final_df == "?")
# Number of nulls per column
colSums(rating_final_df == "?")
```

#### Conclusion
```{r}
sprintf("There are %d unique users.", length(unique(userprofile_df$userID)))
sprintf("There are %d unique restaurants", length(unique(geoplaces2_df$placeID)))
```

## Data Preparation

#### chefmozaccepts.csv

Treatments to be applied to this dataset:
1. Rpayment should be categorized
```{r}
unique(chefmozaccepts_df$Rpayment)

# transform payments to lowercase
chefmozaccepts_df$Rpayment <- sapply(chefmozaccepts_df$Rpayment, tolower)

# We have 615 unique placeID, but only 500 accept cash
unique_rpayment <- unique(chefmozaccepts_df$placeID)
length(unique_rpayment)

# We will categorize Rpayment in 2 categories:
# cash - 0/1
# other - 0/1
chefmozaccepts_df2 <- chefmozaccepts_df
chefmozaccepts_df2$Rpayment_cash <- NA
chefmozaccepts_df2$Rpayment_noCash <- NA
# Remove duplicates based on chefmozaccepts_df.Rpayment column
chefmozaccepts_df2 <- chefmozaccepts_df2[!duplicated(chefmozaccepts_df2$placeID), ]

for(placeID in unique(chefmozaccepts_df$placeID)) {
  payments <- chefmozaccepts_df[which(chefmozaccepts_df$placeID == placeID),]
  if(is.element("cash", payments$Rpayment) & length(payments$Rpayment) > 1) {
    # has both tyoe of payments
    chefmozaccepts_df2[which(chefmozaccepts_df2$placeID == placeID),]$Rpayment_cash <- 1
    chefmozaccepts_df2[which(chefmozaccepts_df2$placeID == placeID),]$Rpayment_noCash <- 1
  }
  else if(is.element("cash", payments$Rpayment) & length(payments$Rpayment) == 1) {
    # only cash
    chefmozaccepts_df2[which(chefmozaccepts_df2$placeID == placeID),]$Rpayment_cash <- 1
    chefmozaccepts_df2[which(chefmozaccepts_df2$placeID == placeID),]$Rpayment_noCash <- 0
  }
  else { # if(!is.element("cash", payments$Rpayment))
    # no cash
    chefmozaccepts_df2[which(chefmozaccepts_df2$placeID == placeID),]$Rpayment_cash <- 0
    chefmozaccepts_df2[which(chefmozaccepts_df2$placeID == placeID),]$Rpayment_noCash <- 1
  }
}

# remove RPayment column because it's no longer necessary
chefmozaccepts_df2 <- subset(chefmozaccepts_df2, select = c("placeID", "Rpayment_cash", "Rpayment_noCash"))
# or chefmozaccepts_df2 <- chefmozaccepts_df2[,-2]
```
Currently Rpayment has 12 possible values: 'cash', 'VISA', 'MasterCard-Eurocard', 'American_Express', 'bank_debit_cards', 'checks', 'Discover', 'Carte_Blanche', Diners_Club', 'visa', 'Japan_Credit_Bureau' and 'gift_certificates'.
We will merge some of them to create the following categories: 
- Cash ('cash' and 'other', because userpayment_df doesn't have 'other' and we assume that 'cash' is a standard payment)
- Debit ('bank_debit_cards')
- Credit ('VISA', 'mastercard-eurocard', 'visa', 'american_express').

#### chefmozcuisine.csv

Treatments to be applied:
1. Rcuisine should be categorized:
  Currently Rcuisine has 7 categories

#### userpayment.csv

```{r}
# We will categorize Rpayment in 2 categories:
# cash - 0/1
# other - 0/1
userpayment_df2 <- userpayment_df
userpayment_df2$Upayment_cash <- NA
userpayment_df$Upayment_noCash <- NA
# Remove duplicates based on chefmozaccepts_df.Rpayment column
userpayment_df2 <- userpayment_df2[!duplicated(userpayment_df2$userID), ]

for(userID in unique(userpayment_df$userID)) {
  payments <- userpayment_df[which(userpayment_df$userID == userID),]
  if(is.element("cash", payments$Upayment) & length(payments$Upayment) > 1) {
    # has both tyoe of payments
    userpayment_df2[which(userpayment_df2$userID == userID),]$Upayment_cash <- 1
    userpayment_df2[which(userpayment_df2$userID == userID),]$Upayment_noCash <- 1
  }
  else if(is.element("cash", payments$Upayment) & length(payments$Upayment) == 1) {
    # only cash
    userpayment_df2[which(userpayment_df2$userID == userID),]$Upayment_cash <- 1
    userpayment_df2[which(userpayment_df2$userID == userID),]$Upayment_noCash <- 0
  }
  else { # if(!is.element("cash", payments$Rpayment))
    # no cash
    userpayment_df2[which(userpayment_df2$userID == userID),]$Upayment_cash <- 0
    userpayment_df2[which(userpayment_df2$userID == userID),]$Upayment_noCash <- 1
  }
}
```

To use 
```{r}
# Join dataframes
final_df <- merge(rating_final_df, geoplaces2_df, by = "placeID")
final_df <- merge(final_df, chefmozaccepts_df, by = "placeID")
final_df <- merge(final_df, chefmozcuisine_df, by = "placeID")
final_df <- merge(final_df, chefmozhours4_df, by = "placeID")
final_df <- merge(final_df, chefmozparking_df, by = "placeID")
final_df <- merge(final_df, userprofile_df, by = "userID")
final_df <- merge(final_df, usercuisine_df, by = "userID")
final_df <- merge(final_df, userpayment_df, by = "userID")
```

```{r}
# Columns
names(final_df)
# Number of attributes
ncol(final_df)
# Number of instances
nrow(final_df)
# Summary
summary(final_df)
# Head
head(final_df)
# Number of nulls
sum(final_df == "?")
# Number of nulls per column
colSums(final_df == "?")
```

# Bibliography

https://www.kaggle.com/liyenhsu/simple-content-based-recommenders





